Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-14 WSSSPE1_report/WSSSPE1_summary_report.tex
fix some typos

\author{Daniel~S.~Katz$^{(1)}$, Hilmar~Lapp$^{(2)}$, Ketan~Maheshwari$^{(3)}$, Frank~L\"{o}ffler$^{(4)}$, 
Matthew~Turk$^{(5)}$, Marcus~D.~Hanwell$^{(6)}$, Nancy~Wilkins-Diehr$^{(7)}$, James~Hetherington$^{(8)}$, 
James~Howison$^{(9)}$, Sou-Cheng~T.~Choi$^{(10)}$, Shel~Swenson$^{(11)}$,  Bruce~Berriman$^{(12)}$, Andy~R.~Terrel$^{(13)}$, Colin~Venters$^{(14)}$}

@@

The second keynote~\cite{WSSSPE1-keynote2} was delivered by Arfon
Smith of GitHub. Smith's talk started with an example from data reduction in Astronomy,
where he needed to remove interfering effects from the device (XXXX? which device). He
built a ``bad pixel mask,'' and realized that while it was persistent,
there was no way or practice of sharing this with the
data. Consequently many researchers repeated the same calculations. He
estimated that 13 person-years were wasted in this redundant
calculation. (XXXX whole paragraph not clear)

@@

Priem and Piwowar argued that assessing the impact of software work
requires looking beyond publications, including evidence of contribution and impact recorded in social
coding-oriented resources such as Github, and conversations about software
in issue trackers, mailing lists, twitter and
beyond~\cite{Priem_WSSSPE}. In keeping with a principle of
the ``altmetrics'' approach, they advocate that scholars should have
resources that empower them to tell their own stories in the
manner most appropriate for them and their audiences.

@@

Christopherson reports on the application of Steps 1-3 to a
computational modeling framework developed in the 1990s. Step 1 was
a 2-day, in-person specifications meeting and code walkthrough. Step 2 was a
5-day hackathon to develop working code, and Step 3 was a 3-day hackathon to refine
the code based on new requirements. The team worked on small, low-risk units of
code. It was challenging, revealed unanticipated obstacles, programmers had to
work together, and experimentation was encouraged.
%
The paper recommended: start small and gradually building toward more complex
objectives, consistent with Agile development; develop consensus before
coding, by repeating step 1 before all hackathons; ensure newcomers
receive orientation prior to the hackathon, such as a code
walkthrough or system documentation; and co-locate collaborators
whenever feasible.

@@

From more traditional open source development, resources were shared that
developed communities explicitly, such as~\cite{citeulike:7888211} 
and~\cite{Trapani:2011}, focusing on
large-scale projects such as the Ubuntu linux distribution and smaller-scale
volunteer-developed projects like such as ThinkUp, respectively.
The process of open source development, while less explicitly focused on
community building, outlined in~\cite{citeulike:478633} was seen as a valuable
resource, particularly when combined with the management and personal
interaction techniques outlined in~\cite{opac-b1134063}.  Growing diversity in
communities was directly addressed in~\cite{Allsopp:2012}, where experiences
growing the diversity of technical conferences in open source were described.

@@

Finally, it appears that the career options of developers from different kinds
of software projects differ significantly. In order to further investigate the
much discussed issue of career paths for developers of these software, we
conducted a small survey of the use-cases discussed above with respect to the
current status of its core developers. We looked at the workshop submissions
and the original cited papers for the respective software. Where available, we
looked at the webpages and source code of the earliest versions to find the
original developers. We included projects older than 5 years and excluded
relatively new projects such as \toolname{SciPy} and \toolname{numFOCUS} from this
survey. we find that in a majority of scientific software more than one
original developers are still around, in some cases for more than 10 years,
(eg. \toolname{R/qtl}).  Whereas, in the case of utility software projects, it
appears that at least one original developer is still around, often in a
part-time or ``benefactor'' role. In other cases, it appears that the original
developers have went on to assume a larger role, with the same project.

>>>>

\author{Daniel~S.~Katz$^{(1)}$, Hilmar~Lapp$^{(2)}$, Ketan~Maheshwari$^{(3)}$, Frank~Löffler$^{(4)}$, 
Matthew~Turk$^{(5)}$, Marcus~D.~Hanwell$^{(6)}$, Nancy~Wilkins-Diehr$^{(7)}$, James~Hetherington$^{(8)}$, 
James~Howison$^{(9)}$, Sou-Cheng~T.~Choi$^{(10)}$, Shel~Swenson$^{(11)}$,  Bruce~Berriman$^{(12)}$, Andy~R.~Terrel$^{(13)}$, Colin~Venters$^{(14)}$}

@@

The second keynote~\cite{WSSSPE1-keynote2} was delivered by Arfon Smith of
GitHub. Smith's talk started with an example from his own past in data
reduction in Astronomy, where he needed to remove interfering effects from the
experimental apparatuses. He built a ``bad pixel mask,'' and realized that
while it was persistent, there was no way or practice of sharing these data
among the scientists. Consequently many researchers repeated the same
calculations. Smith estimated that 13 person-years were wasted by this
repetition.

@@

Priem and Piwowar argued that assessing the impact of software work requires
looking beyond publications, including evidence of contribution and impact
recorded in social coding-oriented resources such as GitHub, and conversations
about software in issue trackers, mailing lists, twitter and
beyond~\cite{Priem_WSSSPE}. In keeping with a principle of the ``altmetrics''
approach, they advocate that scholars should have resources that empower them
to tell their own stories in the manner most appropriate for them and their
audiences.

@@

Christopherson reports on the application of Steps 1-3 to a computational
modeling framework developed in the 1990s. Step 1 was a 2-day, in-person
specifications meeting and code walk-through. Step 2 was a 5-day hackathon to
develop working code, and Step 3 was a 3-day hackathon to refine the code based
on new requirements. The team worked on small, low-risk units of code. It was
challenging, revealed unanticipated obstacles, programmers had to work
together, and experimentation was encouraged.
%
The paper recommended: start small and gradually building toward more complex
objectives, consistent with Agile development; develop consensus before coding,
by repeating step 1 before all hackathons; ensure newcomers receive orientation
prior to the hackathon, such as a code walk-through or system documentation;
and co-locate collaborators whenever feasible.

@@

From more traditional open source development, resources were shared that
developed communities explicitly, such as~\cite{citeulike:7888211}
and~\cite{Trapani:2011}, focusing on large-scale projects such as the Ubuntu
Linux distribution and smaller-scale volunteer-developed projects like such as
ThinkUp, respectively. The process of open source development, while less
explicitly focused on community building, outlined in~\cite{citeulike:478633}
was seen as a valuable resource, particularly when combined with the management
and personal interaction techniques outlined in~\cite{opac-b1134063}.  Growing
diversity in communities was directly addressed in~\cite{Allsopp:2012}, where
experiences growing the diversity of technical conferences in open source were
described.

@@

Finally, it appears that the career options of developers from different kinds
of software projects differ significantly. In order to further investigate the
much discussed issue of career paths for developers of these software, we
conducted a small survey of the use-cases discussed above with respect to the
current status of its core developers. We looked at the workshop submissions
and the original cited papers for the respective software. Where available, we
looked at the web pages and source code of the earliest versions to find the
original developers. We included projects older than 5 years and excluded
relatively new projects such as \toolname{SciPy} and \toolname{numFOCUS} from
this survey. we find that in a majority of scientific software more than one
original developers are still around, in some cases for more than 10 years,
(eg. \toolname{R/qtl}).  Whereas, in the case of utility software projects, it
appears that at least one original developer is still around, often in a
part-time or ``benefactor'' role. In other cases, it appears that the original
developers have went on to assume a larger role, with the same project.
